---
title: "p8105_dmh2228_hw1"
author: "Diana Hernandez"
date: "2023-09-19"
output: github_document
---

## Problem 1 and 2: load tidyverse

```{r}
library(tidyverse)
```

## Problem 1: load moderndive

```{r}
library(moderndive)
```

## Problem 1: load "early_january_weather"

```{r}
data("early_january_weather")
```

## Problem 1: descriptive information of "early_january_weather"

```{r, eval = FALSE}
view(early_january_weather)

head(early_january_weather)

tail(early_january_weather)
```

"early_january_weather" has 358 observations and 15 variables (origin, year, month, day, hour, temp, dewp, humid, wind_dir, wind_speed, precip, pressure, visib, time_hour).

## Problem 1: rows and columnns of "early_january_weather"

```{r}
nrow(early_january_weather)

ncol(early_january_weather)
```

"early_january_weather" has 358 rows and 15 columns. 

## Problem 1: mean temperature in "early_january_weather"

```{r}
mean(early_january_weather$temp)
```

The mean temperature in "early_january_weather" is 39.58212.

## Problem 1: scatter plot of temp and time_hour in "early_january_weather"

```{r}
plot_df = tibble(
  x = early_january_weather$time_hour,
  y = early_january_weather$temp
)

ggplot(plot_df, aes(x = x, y = y, color = early_january_weather$humid)) + geom_point()
ggsave("early_january_weather_plot.pdf", height = 5, width = 5)
```

The scatter plot shows fluctuations between temperature and hour. Humidity varies by day.

## Problem 2: data frame 

```{r}
random_sample_df = tibble(
  norm_samp = rnorm(10),
  vec_logical = norm_samp > 0,
  vec_char = c("This", "is", "an", "example", "of", "using", "data", "frames", "and", "vectors"),
  vec_factor = factor(c("low", "medium", "high", "high", "high", "low", "medium", "low", "high", "low")),
  vec_levels = factor(vec_factor, levels = c("low", "medium", "high"))
)
```

## Problem 2: mean of variables in data frame

```{r}
mean_samp_norm = mean(pull(random_sample_df, norm_samp))
```

```{r}
mean_samp_logic = mean(pull(random_sample_df, vec_logical))
```

```{r}
mean_samp_char = mean(pull(random_sample_df, vec_char))
```

There is no mean for vec_char.

```{r}
mean_samp_factor = mean(pull(random_sample_df, vec_factor))
```

There is no mean for vec_factor.

## Problem 2: using as.numeric 

```{r, eval = FALSE}
num_logical = as.numeric(pull(random_sample_df, vec_logical))
```

```{r, eval = FALSE}
num_char = as.numeric(pull(random_sample_df, vec_char))
```

```{r, eval = FALSE}
num_factor = as.numeric(pull(random_sample_df, vec_factor))
```

as.numeric works on the logical and factor variables, but not the character variable. For the logical and factor variables, R is able to assign numbers to each case because there are groups. The character variable does not have anything to be classified by. This helps explain why a mean is obtained for the logical variable and not the character variable.
